---
title: "Trends in Gm, rwp & cell wall chemistry"
output: html_document
---

Anslysis objective: To see if there is realtionship between cell wall chemistry and cell wall resistance (rwp or Resistwp25) or Gm (conductance).

Problems: 1) the cell wall chemistry was taken from trunk material while Rm and rwp are calculated from leaves. 2) cell wall data were collected from various outside commongarden sites while leaf physiology was collected from greenhouse material. 


```{r, results='hide'}
setwd("~/Desktop/poplar_data")
library(dplyr)

dat1<-read.csv("datainput_2_27_2015.csv")
head(dat1)
names(dat1)

dat1$id<-toupper(dat1$id) #all letters upper case
dat1$id<-gsub("*.CSV", "", dat1$id) # replace .csv to blank
head(dat1)
str(dat1)

datRecal<-read.csv("CLA_Recalci_Input.csv")
datCorv<-read.csv("Corvallis_2010.csv")
datPop<-read.csv("Poplar_C5_C6_12_20_11.csv")

```

getting counts for number of rch and rwp > 0

```{r, eval=FALSE}
a<-dat1 %>% select(id, Resistch25, Resistwp25) %>% filter(Resistwp25 > 0)
b<-dat1 %>% select(id, Resistch25, Resistwp25) %>% filter(Resistch25 > 0)
```

Data manipulation and Gm calculation
Note: can only caluclate Gm for samples that have rwp > 0. 

```{r}
dat2<-dat1 %>% select(id, Duplicate, Jmax25, Rdlight25, Resistwp25, Resistch25, gamma.25, Asat_ALight) %>% filter(Duplicate == "Unique" & Resistwp25 > 0) %>% mutate(Gm = 1/Resistwp25)
```

Investigate relationships or rwp (Resistwp25) with cell wall chemistry

File names for the wall chemistry:
CLA_SG_Input.csv
CLA_C5_6.csv
CLA_Recalci_Input.csv
Corvallis_2010.csv

###Plots for the Clatskine Site

```{r}

datRecal_join<-inner_join(dat2, datRecal, by = "id")
plot(datRecal_join$Lignin, datRecal_join$Resistwp25)
cor.test(datRecal_join$Lignin, datRecal_join$Resistwp25, method = "spearm")
cor.test(datRecal_join$Lignin, datRecal_join$Resistwp25, method = "pearson")
plot(datRecal_join$Lignin, datRecal_join$Gm, ylim=c(0,2))
cor.test(datRecal_join$Lignin, datRecal_join$Gm, method = "spearm")
cor.test(datRecal_join$Lignin, datRecal_join$Gm, method = "pearson")
plot(datRecal_join$SG, datRecal_join$Resistwp25)
plot(datRecal_join$SG, datRecal_join$Gm, ylim=c(0,2))
plot(datRecal_join$Glucose, datRecal_join$Resistwp25)
plot(datRecal_join$Glucose, datRecal_join$Gm, ylim=c(0,2))
plot(datRecal_join$Xylose, datRecal_join$Resistwp25)
plot(datRecal_join$Xylose, datRecal_join$Gm, ylim=c(0,2))
plot(datRecal_join$ GlucoseXylose, datRecal_join$Resistwp25)
plot(datRecal_join$ GlucoseXylose, datRecal_join$Gm, ylim=c(0,2))

```


###Plots for Corvallis
Note, not many points in Corvallis_2010.csv
```{r}

datCorv_join<-inner_join(dat2, datCorv, by = "id")
plot(datCorv_join$Lignin, datCorv_join$Resistwp25)
cor.test(datCorv_join$Lignin, datCorv_join$Resistwp25, method = "spearm")
cor.test(datCorv_join$Lignin, datCorv_join$Resistwp25, method = "pearson")
plot(datCorv_join$Lignin, datCorv_join$Gm, ylim=c(0,2))
cor.test(datCorv_join$Lignin, datCorv_join$Gm, method = "spearm")
cor.test(datCorv_join$Lignin, datCorv_join$Gm, method = "pearson")
plot(datCorv_join$S_G, datCorv_join$Resistwp25)
plot(datCorv_join$S_G, datCorv_join$Gm, ylim=c(0,2))
plot(datCorv_join$Glucoserelease, datCorv_join$Resistwp25)
plot(datCorv_join$Glucoserelease, datCorv_join$Gm, ylim=c(0,2))
plot(datCorv_join$Xyloserelease, datCorv_join$Resistwp25)
plot(datCorv_join$Xyloserelease, datCorv_join$Gm, ylim=c(0,2))
plot(datCorv_join$GlucoseXyloserelease, datCorv_join$Resistwp25)
plot(datCorv_join$GlucoseXyloserelease, datCorv_join$Gm, ylim=c(0,2))
```

### the Poplar_C5_C6_12_20_11.csv dataset
details unknown
```{r}

datPop_join<-inner_join(dat2, datPop, by = "id")
plot(datPop_join$Lignin, datPop_join$Resistwp25)
cor.test(datPop_join$Lignin, datPop_join$Resistwp25, method = "spearm")
cor.test(datPop_join$Lignin, datPop_join$Resistwp25, method = "pearson")
plot(datPop_join$Lignin, datPop_join$Gm, ylim=c(0,2))
cor.test(datPop_join$Lignin, datPop_join$Gm, method = "spearm")
cor.test(datPop_join$Lignin, datPop_join$Gm, method = "pearson")
plot(datPop_join$S_G, datPop_join$Resistwp25)
plot(datPop_join$S_G, datPop_join$Gm, ylim=c(0,2))
plot(datPop_join$C5, datPop_join$Resistwp25)
plot(datPop_join$C5, datPop_join$Gm, ylim=c(0,2))
plot(datPop_join$C6, datPop_join$Resistwp25)
plot(datPop_join$C6, datPop_join$Gm, ylim=c(0,2))

```

**evaluate through correlagrams**
**data from Poplar_C5_C6_12_20_11.csv**

```{r, echo=FALSE, fig.width=10, fig.height=7}
library(corrgram)
corrgram(datPop_join, order=TRUE, lower.panel=panel.shade,
  upper.panel=panel.pie, text.panel=panel.txt,
  main="Populus_tri Poplar_C5_C6_12_20_11") 

```

**data from Corvallis_2010.csv**

```{r, echo=FALSE, fig.width=10, fig.height=7}

corrgram(datCorv_join, order=TRUE, lower.panel=panel.shade,
  upper.panel=panel.pie, text.panel=panel.txt,
  main="Populus_tri phenotype Corvallis_2010") 

```

**data from Clatskine Site **

```{r, echo=FALSE, fig.width=10, fig.height=7}

corrgram(datRecal_join, order=TRUE, lower.panel=panel.shade,
  upper.panel=panel.pie, text.panel=panel.txt,
  main="Populus_tri phenotype Clatskine Site") 
```


