---
title: "Phenotype_exploration"
output: html_document
---

Duplicate analyses were taken for same leaf and clones with the same leaf. we need to see how close these replicates were:

The duplicate analysis needs to be more complete...

```{r, echo=FALSE, results='hide'}
setwd("~/GitHub/poplar_GWAS_photosyn")
library(dplyr)

data1<-read.csv("unique_dup.csv", stringsAsFactors= FALSE)
data2<-read.csv("clone_dup.csv", stringsAsFactors=FALSE)
head(data1)
tail(data1)
head(data2)
tail(data2)

str(data1)

dim(data1)
dim(data2)

A<-data1[2:40, 2:27]
B<-data2[2:40, 2:27]

x<-diag(cor(A[sapply(A, is.numeric)], B[sapply(B, is.numeric)]))
```

ID cleaning start of phenotypic variation analysis:
Note on input data: Need to check original parsing script. Some variables were averaged!!!!

```{r, echo=FALSE}
dat1<-read.csv("fluorescence_formatted_nov26_2014_rename_SLA.csv")
head(dat1)
names(dat1)

dat1$id<-toupper(dat1$id) #all letters upper case
dat1$id<-gsub("*.CSV", "", dat1$id) # replace .csv to blank
head(dat1)
str(dat1)

```

Calculate 'genetic variation,' from Junfei Gu et al. Plant, Cell and Environment (2014) 37, 22–34; which is really phenotypic varaition. According to the authors, genetic variation was calculated as (Xmax − Xmin )/Xmean, where Xmax and Xmin stands for maximum and minimum value in the population, respectively.

```{r}




```

Heat maps for each trait vaition for each individual: calculated as in Junfei Gu et al PC&E 2014. Where (Xi - Xmean)/Xmean, where Xi is the ith genotype and Xmean is the population mean. 

```{r}

jmean<-63.27
jmax.variation<-dat1 %>%
    select(id, Jmax25) %>%
    mutate(Jmax_variation = (Jmax25 - jmean)/jmean)
    
library(ggvis)
 plot(jmax.variation$Jmax25, jmax.variation$Jmax_variation)
plot(jmax.variation$Jmax_variation, jmax.variation2$Jmax_variation)


```









