---
title: "Populus tri photosynthesis phenotype exploration"
output:
  html_document:
    highlight: tango
---

Duplicate analyses were taken for same leaf and clones with the same leaf. we need to see how close these replicates were:

The duplicate analysis needs to be more complete...

```{r, echo=FALSE, results='hide'}
setwd("~/GitHub/poplar_GWAS_photosyn")
library(dplyr)

data1<-read.csv("unique_dup.csv", stringsAsFactors= FALSE)
data2<-read.csv("clone_dup.csv", stringsAsFactors=FALSE)
head(data1)
tail(data1)
head(data2)
tail(data2)

str(data1)

dim(data1)
dim(data2)

A<-data1[2:40, 2:27]
B<-data2[2:40, 2:27]

x<-diag(cor(A[sapply(A, is.numeric)], B[sapply(B, is.numeric)]))
```



```{r, results='hide'}
dat1<-read.csv("datainput_2_27_2015.csv")
head(dat1)
names(dat1)

dat1$id<-toupper(dat1$id) #all letters upper case
dat1$id<-gsub("*.CSV", "", dat1$id) # replace .csv to blank
head(dat1)
str(dat1)

```

Histograms of phenotype/trait/madel parameter distributions:

```{r, echo=FALSE}
library(reshape)
library(ggplot2)
d <- melt(dat1[, c(4:9)])
ggplot(d,aes(x = value)) + 
    facet_wrap(~variable,scales = "free_x") + 
    geom_histogram()

```

Same, but with data filtered Asat < 40 (this excludes 7 observations)

```{r, echo=FALSE}
dat1_clean<-dat1 %>% filter(Asat_ALight < 40)
library(reshape)
library(ggplot2)
d <- melt(dat1_clean[, c(4:9)])
ggplot(d,aes(x = value)) + 
    facet_wrap(~variable,scales = "free_x") + 
    geom_histogram()

```


Histograms of phenotype/trait/madel parameter distributions:

```{r, echo=FALSE}
d <- melt(dat1[, c(10, 12:15)])
ggplot(d,aes(x = value)) + 
    facet_wrap(~variable,scales = "free_x") + 
    geom_histogram()

```

Same, but with data filtered Asat < 40 (this excludes 7 observations)

```{r, echo=FALSE}

d <- melt(dat1_clean[, c(10, 12:15)])
ggplot(d,aes(x = value)) + 
    facet_wrap(~variable,scales = "free_x") + 
    geom_histogram()

```


Histograms of phenotype/trait/madel parameter distributions:

```{r, echo=FALSE}
d <- melt(dat1[, c(20:24, 26)])
ggplot(d,aes(x = value)) + 
    facet_wrap(~variable,scales = "free_x") + 
    geom_histogram()

```

Same, but with data filtered Asat < 40 (this excludes 7 observations)

```{r, echo=FALSE}

d <- melt(dat1_clean[, c(20:24, 26)])
ggplot(d,aes(x = value)) + 
    facet_wrap(~variable,scales = "free_x") + 
    geom_histogram()

```

Histograms of phenotype/trait/madel parameter distributions:

```{r, echo=FALSE}
d <- melt(dat1[, c(27:28)])
ggplot(d,aes(x = value)) + 
    facet_wrap(~variable,scales = "free_x") + 
    geom_histogram()

```

Same, but with data filtered Asat < 40 (this excludes 7 observations)

```{r, echo=FALSE}

d <- melt(dat1_clean[, c(27:28)])
ggplot(d,aes(x = value)) + 
    facet_wrap(~variable,scales = "free_x") + 
    geom_histogram()

```

now phenotype correlations between all variables

```{r, results='hide'}
library(dplyr)
d<-dat1%>%
    select(diam = Av_Diameter_mm, LA=Leaf_Area_cm2, height = Height_cm, SLA = Av_SLA._mg.cm2, Asat = Asat_ALight, Jmax = Jmax25, PSII_effic = PhiPSIIsat_ALight, PSII_dark_effic = PhiPSIILightIntercep, PSI_PSII = betaPSII, CO2_comp = gamma.25, day_resp = Rdlight25, GmRwp = Resistwp25, GmRcp = Resistch25) %>%
    filter(Asat < 40)
   
cor(d, use="complete.obs")
```

*Corrlegrams*

Paramters for the correlagrams:

order=TRUE will cause the variables to be ordered using principal component analysis of the correlation matrix. 

panel.pie (the filled portion of the pie indicates the magnitude of the correlation)

panel.shade (the depth of the shading indicates the magnitude of the correlation)

```{r, fig.width=10, fig.height=7}
library(corrgram)
corrgram(d, order=TRUE, lower.panel=panel.shade,
  upper.panel=panel.pie, text.panel=panel.txt,
  main="Populus_tri phenotype") 
```

Corrlegram with dotplots

```{r, fig.width=10, fig.height=7}
corrgram(d, order=TRUE, lower.panel=panel.ellipse,
  upper.panel=panel.pts, text.panel=panel.txt,
  diag.panel=panel.minmax,
  main="Populus_tri phenotype") 

```

Calculate 'genetic variation,' from Junfei Gu et al. Plant, Cell and Environment (2014) 37, 22–34; which is really phenotypic varaition. According to the authors, genetic variation was calculated as (Xmax − Xmin )/Xmean, where Xmax and Xmin stands for maximum and minimum value in the population, respectively.

```{r, echo=FALSE}
pheno_var<-function(x) {
    a<-(max(x, na.rm=TRUE) - min(x, na.rm=TRUE))/mean(x, na.rm=TRUE)
    a    
}
  
a<-dat1 %>%
    select(diam = Av_Diameter_mm, LA=Leaf_Area_cm2, height = Height_cm, SLA = Av_SLA._mg.cm2, Asat = Asat_ALight, Jmax = Jmax25, PSII_effic = PhiPSIIsat_ALight) %>%
    filter(Asat < 40) %>%
    summarise_each(funs(pheno_var))
library(knitr)
kable(a, digits=2)


b<-dat1 %>%
    select(PSII_dark_effic = PhiPSIILightIntercep, PSI_PSII = betaPSII, CO2_comp = gamma.25, day_resp = Rdlight25, GmRwp = Resistwp25, GmRcp = Resistch25, Asat = Asat_ALight) %>%
    filter(Asat < 40) %>%
    summarise_each(funs(pheno_var))

kable(b, digits=2)

```

distribution of Gm for cell wall and plastid membrane

```{r, echo=FALSE}
library(ggvis)
dat1 %>% filter(Asat_ALight < 40) %>% ggvis(~Resistwp25) %>% layer_histograms()

dat1 %>% filter(Asat_ALight < 40) %>% ggvis(~Resistch25) %>% layer_histograms()

dat1 %>% filter(Asat_ALight < 40) %>% ggvis(~Resistwp25, ~Jmax25) %>% 
    layer_points() %>%
    scale_numeric("y", domain = c(0, 200), nice = FALSE, clamp = TRUE)

dat1 %>% filter(Asat_ALight < 40) %>% ggvis(~Resistwp25, ~betaPSII) %>% layer_points()

dat1 %>% filter(Asat_ALight < 40) %>% ggvis(~Resistwp25, ~gamma.25) %>% 
    layer_points()

dat1 %>% filter(Asat_ALight < 40) %>% ggvis(~Resistwp25, ~Asat_ALight) %>% layer_points()%>%
    scale_numeric("y", domain = c(0, 22), nice = FALSE, clamp = TRUE)
 

```


Heat maps for each trait vaition for each individual: calculated as in Junfei Gu et al PC&E 2014. Where (Xi - Xmean)/Xmean, where Xi is the ith genotype and Xmean is the population mean. 

```{r, echo=FALSE}

dat1 %>%
    select(diam = Av_Diameter_mm, LA=Leaf_Area_cm2, height = Height_cm, SLA = Av_SLA._mg.cm2, Asat = Asat_ALight, Jmax = Jmax25, PSII_effic = PhiPSIIsat_ALight, PSII_dark_effic = PhiPSIILightIntercep, PSI_PSII = betaPSII, CO2_comp = gamma.25, day_resp = Rdlight25, GmRwp = Resistwp25, GmRcp = Resistch25) %>%
    filter(Asat < 40, CO2_comp < 6) %>%
    summarise_each(funs(mean(.,na.rm=TRUE))) %>%
    kable(, digits=2)

ind_phenot_var <- dat1 %>%
        select(ID = num_id, diam = Av_Diameter_mm, LA=Leaf_Area_cm2, height = Height_cm, SLA = Av_SLA._mg.cm2, Asat = Asat_ALight, Jmax = Jmax25, PSII_effic = PhiPSIIsat_ALight, PSII_dark_effic = PhiPSIILightIntercep, PSI_PSII = betaPSII, CO2_comp = gamma.25, day_resp = Rdlight25, GmRwp = Resistwp25, GmRcp = Resistch25) %>%
    filter(Asat < 40) %>%
    mutate(diam_pv = (diam - 10.57)/10.57, SLA_pv = (SLA - 4.19)/4.19, height_pv = (height - 205.85)/205.85, LA_pv = (LA - 63.58)/63.58, Asat_pv = (Asat - 8.56)/8.56, Jmax_pv = (Jmax - 63.98)/63.98, CO2_comp_pv = (CO2_comp - 0.96)/0.96, PSII_effic_pv = (PSII_effic + 2.29)/2.29, PSII_dark_effic_pv = (PSII_dark_effic - 3.94)/3.94, PSI_PSII_pv = (PSI_PSII - 0.35)/0.35, day_resp_pv = (day_resp - 1.51)/1.51, GmRwp_pv = (GmRwp - 2.58)/2.58, GmRcp_pv = (GmRcp - 0.37)/0.37)
    
```

Trait clustering and heat map of the 400 individuals. Each rectangle represents (Xi - Xmean)/Xmean from Junfei Gu et al PC&E 2014, which is an estimate of the phenotypic variation for that trait.

NAs were transformed to 0, which is the mean value for (Xi - Xmean)/Xmean.
```{r, fig.width=10, fig.height=8, echo=FALSE}
# get rid of NAs
# dim(ind_phenot_var) before na removal [416  29]
ind_phenot_var[is.na(ind_phenot_var)] <- 0

input_data <- ind_phenot_var %>% select(ID, diam_pv:GmRcp_pv)

# dim(ind_phenot_var) same as after replacing with 0. The mean for each is 0, so this is cool for heat map representation
library(gplots)
library("RColorBrewer")

#We’ll have to strip off the sample ids and convert them to row names so that the data matrix contains only sequence count data.

row.names(input_data) <- input_data$ID
input_data <- input_data[, -1]

#removing some columns
input_data["GmRcp_pv"]<-NULL
input_data["PSII_effic_pv"]<-NULL
input_data["PSI_PSII_pv"]<-NULL



#basic heatmap (not very informative)
#heatmap(as.matrix(input_data), Rowv = NA, Colv = NA, margins = c(10, 2))

input_data.dist<-dist(input_data)
row.clus <- hclust(input_data.dist, "complete")

# now add row and column -- must transform
input_data.dist.col<-dist(t(input_data))
col.clus <- hclust(input_data.dist.col, "complete")

heatmap.2(as.matrix(input_data), Rowv = as.dendrogram(row.clus), Colv = as.dendrogram(col.clus), margins = c(10, 3), trace = "none", col=greenred(100), breaks=seq(-2.5,2.5,length.out=101))


```

Relationship between Gm and Asat and Jmax

note that photosynthesis values seem low but are on PAR with article: Comparative resource-use efficiencies and growth of Populus trichocarpa and Populus balsamifera under glasshouse conditions1 Virginie M. Pointeau and Robert D. Guy Botany 92: 443–451



