---
title: "Trends in growth, rwp and photosyn"
output: html_document
---

**Objective**: To see if CO2 mesophyll resistance (rwp) is realted to total growth traits.

**Data limiations**: photosynthesis realted traits are from greenhouse and growth characteristics are from field. 

**Growth data files**: Corv_2012_Growth_rep1.csv, Corv_2012_Growth_rep2.csv, Corv_2012_Growth_rep3.csv GWAS_Clatskanis_july_2013_Height_Diameter.csv

**Photosyn data files**:

**Local Git directory**: setwd("~/GitHub/poplar_GWAS_photosyn")


```{r}
setwd("~/GitHub/poplar_GWAS_photosyn")
library(dplyr)

dat1<-read.csv("datainput_2_27_2015.csv")
head(dat1)
names(dat1)

dat1$id<-toupper(dat1$id) #all letters upper case
dat1$id<-gsub("*.CSV", "", dat1$id) # replace .csv to blank
head(dat1)
str(dat1)
```

**Field Traits**    				
HBS3	Height to the bud scar of year 3 (2011 height)				
HBS2	Height to the bud scar of year 2 (2010 height)				
HHB	height to the highest branch			columns F-M are mostly repeated, but with 4 noted trees.	What are they for?
D20	diameter 20cm				
D50	diameter 50cm - for stem taper				
BA	Branch angle				
SB	sylleptic branches from last year's growth (2011)				
LENT	# lenticels in 5cm along trunk				
Tephrina	0-4: 0=zero; 1=1-25%; 2=26-50%; 3=51-75%; 4=76-100%				
Venturia	0-4: 0=zero; 1=1-25%; 2=26-50%; 3=51-75%; 4=76-100%				

```{r}

```

